\documentclass[12pt]{article}
\input{physics1}
\begin{document}

\section*{NYU Physics I---Problem Set 8}

Due Thursday 2016 November 03 at the beginning of lecture.

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
A very thin ladder of length $L$ and mass $M$ leans against a vertical
wall, on a horizontal floor, making an angle of $\theta$ with respect
to the wall.  Imagine that there is a large coefficient of friction
$\mu$ at the floor so that the ladder is in static
equilibrium, but assume that the wall is effectively frictionless.

\textsl{(a)} Draw a free-body diagram for the ladder, showing all
forces acting.

\textsl{(b)} Using the bottom of the ladder as the axis of rotation or
origin, compute all the forces and torques on the ladder such that it
is in equilibrium.

\textsl{(c)} Why did I make the wall ``effectively frictionless''?

\textsl{(d)} Re-solve the problem using the \emph{top} of the ladder
as the axis of rotation or origin.  What is different in the end?

\textsl{(e)} At what angles $\theta$ would the ladder start to slip?
If $\mu=0.8$ (not unreasonable for a ladder with hard rubber feet on a wood
floor), what is the maximum angle at which you could lean the ladder?

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
A long, thin rod of length $L$ and cross-sectional area $A$ and
elastic (Young's) modulus $E$ has mass $M$.

\textsl{(a)} Think of the rod as being like a Hooke's Law spring; it
can be stretched by applying a force.  What is the spring constant $k$
for this spring?

\textsl{(b)} By dimensional analysis, can you combine $L$, $A$, $E$,
and $M$ into a frequency $\omega$?  Do you have more than one choice?  If so,
which of the choices makes most sense? That is, think about how your
answer should scale with changes to the problem.

\textsl{(c)} Look up the properties of a femur bone and compute this
the constant $k$ and the frequency $\omega$ for the femur bone, and
then for a steel rod with the same properties, and then for a
(hypothetical) diamond (!) rod.

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
In lecture, you saw something like the damped harmonic oscillator
differential equation
\begin{equation}
m\,\frac{\dd^2 x}{\dd t^2} + c\,\frac{\dd x}{\dd t} + k\,x = 0 \quad ,
\end{equation}
where $m$ is the mass, $c$ is a damping coefficient, and $k$ is a
restoring constant (a spring constant).  Here we are going to show
that
\begin{equation}
x(t) = A\,\e^{-\frac{\gamma}{2}\,t}\,\cos (\omega\,t + \phi)
\end{equation}
can be a solution to the differential equation.

\textsl{(a)} What are the units of $c$, $A$, $\gamma$, and $\phi$?

\textsl{(b)} Take a derivative of $x(t)$ to get $v(t)$. Take another
to get $a(t)$.

\textsl{(c)} If your derivatives are correct, what is the potential
energy and the kinetic energy as a function of time? Recall that
potential energy for a spring is $(1/2)\,k\,x^2$. What is the total
mechanical energy of the system as a function of time? \emph{Hint:}
The sines and cosines should disappear from this total energy
expression!

\textsl{(d)} Now plug your derivatives into the differential equation,
and see if there is a setting of the parameters $\gamma$ and $\omega$
such that the differential equation can be satisfied? \emph{Hint:}
Group sine and cosine terms separately; both sets of terms must sum to
zero for the differential equation to be satisfied. This is related to
the concept of \emph{detailed balance}.

\textsl{(e)} Did you have to assume things about $m, c, k$ to make
your answer work? What things?

\paragraph{Extra Problem (will not be graded for credit):}%
Re-do the previous problem using complex exponentials. That is, assume
\begin{equation}
x(t) = Z\,\e^{\alpha\,t}
\end{equation}
where $Z$ and $\alpha$ are complex numbers. What's different, and
what's the same?

\paragraph{Extra Problem (will not be graded for credit):}%
Show that these two descriptions of a simple harmonic oscillator
\begin{equation}
x(t) = A\,\cos(\omega_0\,t) + B\,\sin(\omega_0\,t)
\end{equation}
\begin{equation}
x(t) = X\,\cos (\omega_0\,t+\phi)
\end{equation}
are completely equivalent by finding the relationship between $A, B$
and $X, \phi$ that makes them identical.

\end{document}
