\documentclass[12pt]{article}
\input{physics1}
\begin{document}

\section*{NYU Physics I---Problem Set 1}

This is due Tuesday 2025 September 9 by noon (2025-09-09T12:00) New York civil time.
Your submitted Problem Set solution must include a ``credits''
section, in which you explicitly give credit to every resource that
you used.  This might include the textbook, online resources, fellow
students, tutors, or anyone or anything else you consulted in completing the
problems.

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
If a car gets 27 miles to the gallon, how many liters of fuel does it
take to go $100\,\km$? (In Europe, fuel efficiencies are given in the
latter terms.)

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
\textsl{(a)}~What is the maximum amount of cash you can obtain by
successfully robbing an armored truck?  Assume that it is packed with
twenties; that is, estimate an answer by considering the volume of the
truck, and (harder to estimate) the volume of a 20-dollar bill.  State
your assumptions and explain your work, but please don't attempt an
experiment.  Be sure to explain exactly how you estimated the volume
of a 20-dollar bill.  \emph{Hint: think of a stack of bills to
  estimate the volume.  Feel free to \emph{check} any part of your
  answer on the internet, but make sure you actually make a justified,
  quantitative estimate independently.}

\textsl{(b)}~Do you think that many of the armored trucks in Manhattan
are fully packed with 20-dollar bills?

\textsl{(c)}~Would a similar truck weigh more, less, or about the same
if it contained the same amount of money but in the form of gold bars
instead of 20-dollar bills?

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
Imagine you have a mass $M$, a length $h$, a velocity $v$, and an
acceleration $g$. What combinations of these can you make that will
have units of
\textsl{(a)}~time,
\textsl{(b)}~force, and
\textsl{(c)}~energy?
Don't try to be exhaustive; just try to get two different expressions
for each!

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
A bit on air resistance and terminal velocities. If you want more
discussion of these issues, see \url{http://arxiv.org/abs/0709.0107}
(click the PDF link at the right of the page).

\textsl{(a)}~Show that the ram-pressure-force formula $\rho\,A\,v^2$ has the
correct units to be a force, when $\rho$ is a mass density (mass per volume),
$A$ is a cross-sectional area, and $v$ is a speed. Look up air drag on
Wikipedia and see what this formula is missing.

\textsl{(b)}~At what downward falling speed $v$ does an object with
mass $M$ and cross-sectional area $A$ find that ram pressure balances
the gravitational force? Derive an expression. This is the formula
(ish) for the terminal velocity!

\textsl{(c)}~Two pennies are dropped (carefully) from a tall building,
one so that it falls precisely edge-on, and one so that it falls
precisely face-on (difficult but not impossible in practice).  What
(roughly) are the two terminal velocities and what (roughly) do you expect to be
their ratio $v_{\mathrm{edge}}/v_{\mathrm{face}}$?

\paragraph{Problem~\theproblem:}\refstepcounter{problem}%
\textsl{(a)}~Roughly how far does a typical American car have to drive
to ``sweep up'' or drive through a column of air that is comparable in
mass to the car itself?  You will have to estimate the
cross-sectional area and mass of a typical car (or look both things
up on the web; if you look them up, give the make and model).

\textsl{(b)}~Describe in words the \emph{environmental significance} of this
distance.
If you want a (wrong but useful) model for how the car interacts with the
air, imagine that the air has to ``bounce off'' the front of the car.
That is, the car has to impart energy to the air.

\end{document}
