\documentclass[12pt]{article} \input{physics1}
\begin{document}

\noindent
Name: \rule[-1ex]{0.55\textwidth}{0.1pt}
NetID: \rule[-1ex]{0.2\textwidth}{0.1pt}

\section*{NYU Physics I---Term Exam 4}

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from lecture 2018-10-30)
What is the pressure difference, roughly, between the top surface
of an ice cube and the bottom, when that ice cube is floating in water?
Imagine that the ice cube is $2\,\cm$ on a side, and make any other
(correct) assumptions you need to make. Give your answer in Pa or $\N\,\m^{-2}$.

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Problem Set 7)
You calculated that a pendulum with a period of $2\,\s$ has a length
very close to $1\,\m$. What would be the length of a pendulum with
a period of $8\,\s$?

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Problem Set 8)
A very thin ladder of length $L$ and mass $M$ leans against a
vertical wall, on a horizontal floor, making an angle of $\theta$ with respect to the
wall. Imagine that there is a large coefficient of friction $\mu$ at the floor such that
the ladder is in static equilibrium, but assume that the wall is effectively
frictionless. Draw a free-body diagram for the ladder, showing all forces acting.

\vfill
~
\clearpage

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Problem Set 8)
In the equation
$$
\frac{\dd^2 x}{\dd t^2} + Q\,\frac{\dd x}{\dd t} + P\,x = 0 \quad ,
$$ what are the units of $Q$ and $P$?

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from worksheet on potentials)
If you have a potential of the form
$$
U(x) = A\,x^3 - B\,x + C
$$ where $A$ and $B$ and $C$ are positive constants, find a location $x=x_0$ where
the force is zero.

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from worksheet on ideal gas)
If you have a number density $n$ (that is, something with units of particles per unit volume)
and they are moving at speed $v$ in the $x$ direction, how many particles hit a wall (wall
oriented perpendicular to the $x$ direction) of area
$L^2$ per unit time? Any dimensionally correct combination of $n$, $v$, and $L$ will be accepted
as correct.

\vfill
~
\end{document}
