\documentclass[12pt]{article}
\input{physics1}
\begin{document}

\noindent
Name: \rule[-1ex]{0.55\textwidth}{0.1pt}
NetID: \rule[-1ex]{0.2\textwidth}{0.1pt}

\section*{NYU Physics I---Term Exam 2}

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Lecture on 2018-09-27)
A roller-coaster cart is at the top of a loop-the-loop (and therefore
upside-down). The trajectory of the center of mass of the cart has a
radius of curvature $R=5\,\m$. How fast does the roller-coaster have
to be moving in $\mps$ to stay on it's proper path (that is, on the
tracks)?  Assume the mass is $M=1000\,\kg$ and the acceleration due to
gravity is $g=10\,\mpss$.

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Lecture on 2018-09-25)
In \emph{16 words or fewer} tell me why the mass flying off the
(not a) aki jump didn't fly
all the way back up to the release height. Put a box around your answer,
so I can count the words!

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Problem Set 3)
If a runner, starting at rest, accelerates at $5\,\mpss$ for $2\,\s$
and then continues at constant speed for $19\,s$ more, how far will
she have run at the end of that $21\,\s$?

\vfill
~

\clearpage
\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from Problem Set 4)
What is your kinetic energy when you are walking along the street?
State your assumptions, and make sure they are \emph{reasonable.}

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from the blocks-and-pulleys worksheet)
A massless pulley hangs from the ceiling from a string which is at
tension $T_1$. Over this pulley is another string at tension $T_2$, on
the ends of which are massive blocks attached. What is the
relationship between $T_1$ and $T_2$? If you have to assume additional
things to solve this problem, state them.

\vfill

\paragraph{\problemname~\theproblem:}\refstepcounter{problem}%
(from the friction worksheet)
You have a block of mass $m$ on an inclined plane, inclined at an
angle $\theta=15\,\deg$ to the horizontal. The coefficient of friction
is $\mu=0.9$. What is the magnitude of the frictional force on the
block? The acceleration due to gravity is $g$.
You can leave your answer in terms of $\mu$, $m$, $g$, $\theta$, or
whatever you need to deliver a correct answer.
Once again, state any assumptions you need to make.

\vfill
~
\end{document}
