\documentclass[12pt]{article}
\input{physics1}
\begin{document}

\section*{NYU Physics I---numerical integration}

The question here is: How long does it take for an object to fall
$10\,\m$ near the surface of the Earth in the absence of air
resistance?  You will need to perform numerical integrations
throughout this course, so we begin by solving the problem that way.

\paragraph{\theproblem}\refstepcounter{problem}%
Choose a partner (if you wish), obtain a sheet of notebook paper and
put your names on it.  Start a table with 21 rows and four columns for
four variables $t$, $a$, $v$, and $x$.

\paragraph{\theproblem}\refstepcounter{problem}%
In the $t$ column, put in times $0.0\,\s$, $0.1\,\s$, $0.2\,\s$,
\ldots, $2.0\,\s$.  In the $a$ column, put $-9.8\,\mpss$ in all 21
rows.

\paragraph{\theproblem}\refstepcounter{problem}%
In the first entry of the $v$ column, put $0.0\,\mps$.  In each
subsequent row $i+1$ put in for the $v$ entry $v_{i+1}$:
\begin{equation}
v_{i+1}= v_i + a_i\,\Delta t = v_i + a_i\,\left[t_{i+1}-t_i\right] \quad .
\end{equation}
Do that for all 20 subsequent rows.  Keep ``sufficient'' accuracy as
you go. Use a phone or calculator; no need to do this calculating by hand!

\paragraph{\theproblem}\refstepcounter{problem}%
In the first entry of the $x$ column, put $10.0\,\m$.  In each
subsequent row $i+1$ put in for the $x$ entry $x_{i+1}$:
\begin{equation}
x_{i+1}= x_i + v_i\,\Delta t = x_i + v_i\,\left[t_{i+1}-t_i\right] \quad .
\end{equation}
Do that for all 20 subsequent rows.  Keep ``sufficient'' accuracy as
you go.

\paragraph{\theproblem}\refstepcounter{problem}%
(Answer this and the following parts on your sheet of paper, below the
table or on the other side.)  Does the object get to $x=0\,\m$ within
the time span in your table?  If it does, make an estimate of the time
$t$ at which the object gets to $x=0\,\m$.

\paragraph{\theproblem}\refstepcounter{problem}%
Do you expect your answer to be an overestimate or underestimate of
the true time obtained by an analytic calculation?  Give a written
explanation.

\paragraph{\theproblem}\refstepcounter{problem}%
Do you recall this equation?
\begin{equation}
x = x_0 + v_0\,t + \frac{1}{2}\,a\,t^2
\end{equation}
Use this to compute the true time it takes the object to fall
$10\,\m$.  Were you correct in your expectation?  How far off was your
numerical answer?

\paragraph{\theproblem}\refstepcounter{problem}%
Name two things you could have done to make your integration more
accurate.

\end{document}
