\documentclass[12pt]{article}
\input{physics1}
\begin{document}

\section*{NYU Physics I---oscillations}

Consider a mass $M$ on a spring of spring constant $k$, released from
rest at $t=0$ but from a distance $X$ (in the $x$-direction, which is
parallel to the spring) away from the equilibrium position for the
mass.  Assume there are no other forces acting!

With a partner, do the following graphing assignments. They might
require some planning, because we want to align all the time axes.  In
all the parts below, when we say ``label'' we mean ``give the
horizontal and vertical position of''. That is, we want the graphs to
be quantitatively correct.

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the position of the mass as a function of time for a few periods.
Label minima, maxima, and zero crossings in time, and label the amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the velocity of the mass as a function of time for the same time
interval, on a new figure, but time-aligned with the previous plot.  Label
minima, maxima, and zero crossings in time, and label the velocity amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the acceleration of the mass as a function of time for the same time
interval, on a new figure, but time-aligned with the previous plot.  Label
minima, maxima, and zero crossings in time, and label the acceleration amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the kinetic energy of the mass as a function of time for the same time
interval, on a new figure, but time-aligned with the previous plot.  Label
minima, maxima, and zero crossings (if any) and label the kinetic-energy amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the kinetic energy of the mass as a function of time for the same time
interval, on a new figure, but time-aligned with the previous plot.  Label
minima, maxima, and zero crossings (if any) and label the kinetic-energy amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Work out the potential energy of a spring stretched by $x$ by
integrating force times distance.

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the potential energy of the mass as a function of time for the same time
interval, on a new figure, but time-aligned with the previous plot.  Label
minima, maxima, and zero crossings (if any) and label the potential-energy amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Guess what you will get if you plot the sum of the potential and kinetic energy plots!

\paragraph{\theproblem}\refstepcounter{problem}%
Plot the total of potential plus kinetic energy as a function of time for the same time
interval, on a new figure, but time-aligned with the previous plot.  Label
minima, maxima, and zero crossings (if any) and label the total-energy amplitude.

\paragraph{\theproblem}\refstepcounter{problem}%
Guess what your plots would have looked like if there had been a bit of
damping---a bit of air drag or friction.

\end{document}
