\documentclass[12pt,letterpaper]{article}
\pagestyle{empty}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\topmargin}{-0.5in}
\setlength{\textheight}{10.0in}
\setlength{\oddsidemargin}{-0.5in}
\setlength{\textwidth}{7.5in}
\begin{document}

A block of mass $5~\mathrm{kg}$ moving initially at speed
$7~\mathrm{m\,s^{-1}}$ in the positive $x$ direction collides
elastically with another block of mass $2~\mathrm{kg}$ moving
initially at speed $2~\mathrm{m\,s^{-1}}$ in the negative $x$
direction.  Assume that the collision happens entirely along the $x$
axis and that there is no friction or anything else.

(a) Before the collision, what is the total momentum (sum of all
momenta) and the center-of-mass velocity?

\vfill

(b) Before the collision, what are the velocities in the
``center-of-mass'' frame (ie, relative to the center of mass)?

\vfill

(c) What are the individual momenta (magnitudes and directions) and
what is the total momentum in the center-of-mass frame, before the
collision?

\vfill

(d) How can you write the kinetic energy ($[1/2]\,m\,v^2$) of a block
in terms of the momentum $p$ and mass $m$?

\vfill

(e) What are the individual momenta and total momenta in the
center-of-mass frame, immediately after the collision?  Assume the
collision happens only along the $x$ direction.

\vfill

(f) How do you transform back to the ``laboratory'' frame (the
original frame of the problem statement)?

\vfill

(g) What are the velocities (magnitudes and directions) of the blocks
in the laboratory frame, after the collision?

\vfill

(h) Compute the total momentum and total kinetic energy, before and
after, in the laboratory frame.  Do they agree?

\vfill

Discuss.

\end{document}
